---
title: "stdmodsem"
author: "Shu Fai Cheung"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stdmodsem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What It Can Do?

## It can

- Compute an unstandardized or standardized indirect effect or
  conditional indirect effect in a path model.
- Form bootstrap confidence for this effect.

## Advantages

- No need to define any parameters when fitting a model. Just fit the model first,
  and then compute the effect for any pathway, from any variable to any other
  variable.
- Supports models fitted by path analysis using `lavaan::sem()` or by
  a set of multiple regression models using `lm()`.
- No limit on the number of predictors, mediators, and outcome variables, other
  than that by `lavaan::sem()`.
- Supports dataset with missing data through `lavaan::sem()`.
- Supports numeric and categorical moderators.
- Bootstrap, which is time consuming, can be conducted just once. Estimating
  the effects and forming their confidence intervals can reuse the precomputed
  bootstrap samples.

## Limitations (For Now)

- Does not support categorical predictors.
- Does not support indirect effects between latent variables.

# Moderated Mediation: Multiple Regression by `lm()`

This is the sample dataset comes with the package:

```{r}
library(stdmodmed)
dat <- modmed_x1m3w4y1
print(head(dat), digits = 3)
```

Suppose this is the model being fitted:

```{r}
lm_m1 <- lm(m1 ~ x * w1, dat)
lm_y <- lm(y ~ m1 * gp + x, dat)
```

This is a mediation model, `x -> m1 -> y`, with both paths
moderated, `x -> m1` by `w1`, and `m1 -> y` by `gp`, `gp`
a categorical variable.

These are the coefficients:

```{r}
printCoefmat(summary(lm_m1)$coefficients, digits = 3)
printCoefmat(summary(lm_y)$coefficients, digits = 3)
```

To compute conditional indirect effects and form bootstrap confidence
intervals, we use `cond_indirect_effects()`. For a model with coefficients
estimate by multiple regression, we first group the `lm()` output to a `list`:

```{r}
fit <- list(lm_m1, lm_y)
```

We then use `mod_levels_list()` to generate combinations of the levels of the
moderators:

```{r}
mod_levels <- mod_levels_list("w1",
                              c("gpmars", "gpvenus"),
                              fit = fit,
                              sd_from_mean = c(-1, 1),
                              merge = TRUE)
mod_levels
```

To form bootstrapping confidence intervals, we first repeat the regression
analysis in all models by nonparametric bootstrapping and store the results.
This can be done by `lm2boot_out()`:

```{r}
fit_boot_out <- lm2boot_out(fit, R = 2000, seed = 87415)
```

It is all set. We can now compute the conditional indirect effect:

```{r}
out <- cond_indirect_effects(wlevels = mod_levels,
                             x = "x",
                             y = "y",
                             m = "m1",
                             fit = fit,
                             boot_ci = TRUE,
                             boot_out = fit_boot_out)
out
```

# Moderated Mediation: Path Analysis by `lavaan::sem()`

Suppose we fit the same model, but by path analysis. `lavaan::sem()` requires
users to create the indicator variables for categorical variables. This can
be done b y `factor2var()`:

```{r}
dat <- modmed_x1m3w4y1
dat <- data.frame(dat,
                  factor2var(dat$gp, prefix = "gp", add_rownames = FALSE))
print(head(dat), digits = 2)
```

We can fit the model now:

```{r}
library(lavaan)
mod <-
"
m1 ~ x + w1 + x:w1
y ~ m1 + gpmars + gpvenus + m1:gpmars + m1:gpvenus + x
"
set.seed(870454)
fit <- sem(model = mod,
           data = dat,
           fixed.x = FALSE,
           se = "boot",
           bootstrap = 2000)
```

```{r}
mod_levels <- mod_levels_list("w1",
                              c("gpmars", "gpvenus"),
                              fit = fit,
                              sd_from_mean = c(-1, 1),
                              merge = TRUE)
mod_levels
```

```{r}
boot_out <- fit2boot_out(fit = fit)
```


```{r}
out <- cond_indirect_effects(wlevels = mod_levels,
                             x = "x",
                             y = "y",
                             m = "m1",
                             fit = fit,
                             boot_ci = TRUE,
                             boot_out = boot_out)
out
```

# Mediation: Multiple Regression by `lm()`

# Mediation: Path Analysis by `lavaan::sem()`

# Further Information