---
title: "Technical Appendix: Workflow of `get_prod()` and Friends"
author: "Shu Fai Cheung & Sing-Hang Cheung"
output:
  html_document:
    fig.align: "center"
    toc: true
    number_sections: false
bibliography: references.bib
csl: apa.csl
---

```{r echo = FALSE}
library(DiagrammeR)
```

# Goal

This technical appendix describes how product terms
in a model or dataset is identified in the package
[manymome](https://cran.r-project.org/package=manymome).

The function, `get_prod()`, is called internally
by other main functions. Users do not need to call
it directly. Nevertheless, advanced users may be
interested in learning how it works.

# `get_prod()`

This function is used to find all product term(s), if any,
from `x` to `y`, `x` having a direct path to `y`.

## Main arguments

- `x`: The name of the predictor.

- `y`: The name of the outcome.

- `fit`: A fit object. Currently only supports a
 `lavaan`-class object.

- `est`: The output of `lavaan::parameterEstimates()`. If
 `NULL`, the default, it will be generated from `fit`.
 If supplied, `fit` will ge ignored.

- `data`: A data frame. If supplied, it will be used
  to identify the product terms.

- `operator`: The string used to indicate
  a product term. Default is `":"`, used in both `lm()`
  and `lavaan::sem()` for observed variables. If raw
  data is not available, this is needed for identifying
  product terms.

## Output

This is a sample output:

```r
$prod
[1] "x_o_w1"

$b
   x_o_w1
0.2337979

$w
[1] "w1"

$x
[1] "x"

$y
[1] "m1"
```

It returns a list with these elements:

  - `prod`: The name(s) of the product term(s) in the model or data.
    It can have more than one element if the effect of `x` on `y`
    is moderated by more than one moderator.

  - `b`: The coefficient(s) of the product term(s).

  - `w`: The name(s) of the moderator(s).

  - `x`: The predictor, `x`.

  - `y`: The outcome variable, `y`.

## Sample Data and Model

The sample dataset `modmed_x1m3w4y1`
from the package `manymome` will be used for
illustration:

```{r}
library(manymome)
dat <- modmed_x1m3w4y1
print(head(dat), digits = 3)
```

For illustration, some product terms will be formed manually,
and some product terms will be formed using the `:` operator
in `lavaan`:

```{r}
dat$x_o_w1 <- dat$x * dat$w1
library(lavaan)
mod <-
"
m1 ~ x   + w1 + x_o_w1
m2 ~ m1  + w2 + m1:w2
m3 ~ m2
y  ~ m3  + w4 + m3:w4 + x + w3 + x:w3 + x:w4
"
fit <- sem(model = mod,
           data = dat,
           meanstructure = TRUE,
           fixed.x = FALSE)
```

Fit the model by `lm()`:

```{r}
lm_m1 <- lm(m1 ~ x*w1, dat)
lm_m2 <- lm(m2 ~ m1*w2, dat)
lm_m3 <- lm(m3 ~ m2, dat)
lm_y <- lm(y ~ m3*w4 + x*w3 + x*w4, dat)
lm_list <- lm2list(lm_m1, lm_m2, lm_m3, lm_y)
lm_est <- manymome:::lm2ptable(lm_list)
```

## Modes

It works in several modes.

### `fit` only, with raw data

A `fit` object (`lavaan`-class object) only is supplied,
which has raw data stored.

The parameter estimate table, `est`, is extracted from `fit`.

Raw data is extracted from `fit` to identify product terms.

It then returns product terms that involve `x` in predicting `y`, if any.

Case: A path with a moderator

```{r}
get_prod(x = "x", y = "m1", fit = fit)
```

Case: A path with more than one moderator

```{r}
get_prod(x = "x", y = "y", fit = fit)
```

Case: A path without a moderator

```{r}
get_prod(x = "m2", y = "m3", fit = fit)
```

Case: `x` does not have a direct path to `y`

```{r}
get_prod(x = "m1", y = "m3", fit = fit)
```

### `est` and `data`

In this mode, raw data supplied directly is used to identify product terms.

The parameter estimate table `est` is used to determine the form of the model.

It then returns product term(s) that involve `x` in predicting `y`, if any.

Case: A path with a moderator

```{r}
get_prod(x = "x", y = "m1", est = lm_est$est, data = lm_est$data)
```

Case: A path with more than one moderator

```{r}
get_prod(x = "x", y = "y", est = lm_est$est, data = lm_est$data)
```

Case: A path without a moderator

```{r}
get_prod(x = "m2", y = "m3", est = lm_est$est, data = lm_est$data)
```

Case: `x` does not have a direct path to `y`

```{r}
get_prod(x = "m1", y = "m3", est = lm_est$est, data = lm_est$data)
```

### `est` only

In this mode, `operator` needs to be set to identify product terms from
the parameter estimate table `est`.

The table from `est` is then used to determine the form of the model.

It then returns product term(s) that involve `x` in predicting `y`, if any.

This mode is useful when a product term is not formed from the raw data. For
example, when the product term is a product of two latent variables, or
when raw data is not available.

Case: A path with a moderator

```{r}
get_prod(x = "x", y = "m1", est = lm_est$est, operator = ":")
```

Case: A path with more than one moderator

```{r}
get_prod(x = "x", y = "y", est = lm_est$est, operator = ":")
```

Case: A path without a moderator

```{r}
get_prod(x = "m2", y = "m3", est = lm_est$est, operator = ":")
```

Case: `x` does not have a direct path to `y`

```{r}
get_prod(x = "m1", y = "m3", est = lm_est$est, operator = ":")
```

## Workflow

```{r fig-get_prod, fig.cap = "Workflow of `manymome::get_prod()`", echo = FALSE}
mermaid("
flowchart TD
  classDef default fill:#EEEEFF;
  classDef errornode fill:#FFDDDD;
  classDef startend fill:#DDFFDD;
  classDef mcnode fill:#FFFFDD;
  classDef bootnode fill:#DDFFFF;
  classDef lavnode fill:#FFDDFF;
  classDef lmnode fill:#FFDDDD;

  ZZ([\"Start\"])
  ZZ:::startend -->

  ZA{{\"est NULL?\"}}
  estNULL[\"Set est by</br>lavaan::parameterEstimates(fit)\"]
  ZA -- Yes --> estNULL
  estNULL --> fittype
  ZA -- No --> fittype

  fittype{{\"fit a lavaan-object?\"}}
  fittype -- lavaan --> findalllavaan
  fittype -- No --> data
  findalllavaan[\"Call find_all_products()</br>to find all columns that are</br>the products of other columns,</br>using data from fit\"]
  data{{\"data supplied?\"}}
  findalllavaan --> data
  findalldata[\"Call find_all_products()</br>to find all columns that are</br>the products of other columns,</br>using data from argument\"]
  data -- Yes --> findalldata
  findalldata --> yadv
  data -- No --> yadv

  yadv{{\"Is y the DV of any path?\"}}
  ynotdv([\"y not a DV. Return NA\"])
  yadv -- No --> ynotdv:::startend
  yadv -- Yes --> allprodsAllNA

  allprodsAllNA{{\"all_prods NA?\"}}
  allprodsAllNA -- No --> iprod
  allprodsAllNA -- Yes --> prodbyop
  iprod[\"Set i_prod_rhs (logical) to</br>path(s) with product term(s)\"]
  prodbyop[\"Find product term(s)</br>by operator</br>and set i_prod_rhs (logical)\"]
  iprod --> iprodAny
  prodbyop --> iprodAny

  iprodAny{{\"Any i_prod_rhs TRUE?\"}}
  iprodAny -- All FALSE --> iprodFALSE:::startend
  iprodFALSE([\"No product term along the path.</br>Return NA\"])
  iprodAny -- At least one TRUE --> allprodsNA

  allprodsNA{{\"all_prods NA?\"}}
  allprodsNA -- No --> xinprodschk

  xinprodschk{{\"x in any product term(s)?\"}}
  xinprodsTRUE[\"Set prod_x to</br>the product term(s) and</br>w to the moderator(s)</br>using all_prods\"]
  xinprodsFALSE[\"Set prod_x and w </br>to NULL\"]
  xinprodschk -- Yes --> xinprodsTRUE
  xinprodschk -- No --> xinprodsFALSE

  allprodsNA -- Yes --> xinprodsop
  xinprodsop[\"Find x in product terms</br>using operator\"]
  xinprodsop --> xinprodsopchk
  xinprodsopchk{{\"x in any product term(s)?\"}}
  xinprodsopTRUE[\"Set prod_x to</br>the product term(s) and</br>w to the moderator(s)</br>using opeator\"]
  xinprodsopFALSE[\"Set prod_x and w </br>toNULL\"]
  xinprodsopchk -- Yes --> xinprodsopTRUE
  xinprodsopchk -- No --> xinprodsopFALSE

  xinprodsTRUE & xinprodsopTRUE & xinprodsFALSE & xinprodsopFALSE --> getb
  getb[\"Use sapply on get_b()</br>to get the coefficient(s) of</br>the product term(s)\"]

  getb --> results:::startend
  results([\"Pack the results and return a list\"])

", height = 2000, width = 800)
```

# `find_all_products()` and `find_product()`

## How it works

The function `find_all_products()` is an internal function
for identifying all product terms in a dataset. It requires
the raw data to work. It is called by `get_prod()` when raw
data is available.

For each column in a dataset, it calls `find_product()` to check
whether this column is the product of two other columns. If yes,
`find_product()` returns the component column names. It then
returns a named list of the output of `find_product()` for
columns of product term.

The search by `find_product()` is done numerically, checking
whether a column can be formed from the product of
any two other columns (missing data allowed).

This approach is exact and simple. There is no need to rely on
any special naming convention in the model to denote a product term.

`find_all_products()` returns a named list of product term components.

If `expand` = TRUE, it will try to find all lowest order components.
Therefore, a product term of three or more columns can also be identified.

## Example

We first create a dataset with some product terms.

```{r}
library(manymome)
set.seed(63224)
dat <- round(as.data.frame(MASS::mvrnorm(100, rep(0, 5), diag(5))), 3)
head(dat)
dat$V2V1 <- dat$V2 * dat$V1
dat$V3V5 <- dat$V3 * dat$V5
dat$V1V2V3 <- dat$V1 * dat$V2 * dat$V3
print(head(dat), digits = 3)
```

```{r}
manymome:::find_all_products(dat)
```

The output is a named list of character vectors. The names are the
column names that are product terms. Each character vector stores the
column names of its components.

